---
title: "Assignment 2: Data transformation"
author: Monal Daterao
date: October 5th, 2023
output: html_document
---

This is a homework assignment I completed that showcases some of my data transformation
skills in R. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(tidyverse)
load("nmes2018.rda")
```

### Question 1

* How many people in the dataset have an MSCD?

Answer:
```{r}
by_mscd <- group_by(nmes.data, mscd)
summarize(by_mscd, n = n())

```

* What was the highest medical expenditure in this data?

Answer:
```{r}
summarize(nmes.data, max_exp = max(totalexp, na.rm=TRUE))

```

* What was the highest medical expenditure for someone without an MSCD?

Answer:
```{r}
by_mscd <- group_by(nmes.data, mscd)
summarize(by_mscd, max_exp = max(totalexp, na.rm=TRUE))

```

* How old is the youngest person with an MSCD?

Answer:
```{r}
by_mscd <- group_by(nmes.data, mscd)
summarize(by_mscd, youngest = min(age, na.rm=TRUE))

```

* How old is the oldest person without an MSCD?

Answer:
```{r}
by_mscd <- group_by(nmes.data, mscd)
summarize(by_mscd, oldest = max(age, na.rm=TRUE))

```


### Question 2
Write R code to reproduce tables that compare people with and without an MSCD for the variables age, bmi, educate, poor, and female.

```{r}
by_mscd <- group_by(nmes.data, mscd)

summarize(by_mscd,
          n = n(),
          mean_age = mean(age, na.rm=TRUE),
          sd_age = sd(age, na.rm=TRUE))

summarize(by_mscd,
          n = sum(!is.na(bmi)),
          mean_bmi = mean(bmi, na.rm=TRUE),
          sd_bmi = sd(bmi, na.rm=TRUE))

nmes.data %>%
  group_by(mscd, educate) %>%
  summarize(n = n()) %>%
  mutate(percent = n/sum(n) * 100)

nmes.data %>%
  group_by(mscd, poor) %>%
  summarize(n = n()) %>%
  mutate(percent = n/sum(n) * 100)

nmes.data %>%
  group_by(mscd, female) %>%
  summarize(n = n()) %>%
  mutate(percent = n/sum(n) * 100)
```

### Question 3

Create boxplots that show the relationship between medical expenditure and MSCD status by working the log10-transformed expenditures rather than medical expenditures on the dollar scale.

Plot 1:
```{r}
nmes.data %>%
  mutate(log10 = log10(totalexp)) %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = mscd, y = log10, fill = mscd), varwidth = TRUE) +
  facet_grid(poor ~ ageCat) +
  labs( title = "Medical expenditures by major smoking−caused disease (MSCD)",
    subtitle = "Stratified by age group and poverty status",
    y = "Total medical expenditures ($)",
    x = "Major smoking-caused disease (MSCD) status", 
    fill = "MSCD status"
    )+
  theme(legend.position = "bottom")

  

```

Plot 2:
```{r}
nmes.data %>%
  mutate(log10 = log10(totalexp),
         agecatpoor = paste(ageCat, poor, sep = " ")) %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = agecatpoor, y = log10, fill = mscd), varwidth = TRUE) +
  labs( title = "Medical expenditures by major smoking−caused disease (MSCD)",
    subtitle = "Stratified by age group and poverty status",
    y = "Total medical expenditures ($)",
    x = "Major smoking-caused disease (MSCD) status", 
    fill = "MSCD status"
    )+
  theme(legend.position = "bottom")

```
