---
title: "Assignment 4: Writing functions"
author: "Monal Daterao"
date: "November 26th, 2023"
output: html_document
---

This is a homework assignment I completed that showcases my ability to write functions in R. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

(1) Write a function called `mean_ci()` for computing 95% confidence intervals for either a single mean or a difference in means.

```{r}
mean_ci <- function(x, y=NULL, type= "one.mean", ...) {
  if(type == "one.mean") {
    if(!is.numeric(x)) {
      stop("Please enter a numeric value for x!")
    }
    xbar <- mean(x, ...)
    n <- sum(!is.na(x))
    se <- sd(x, ...) / sqrt(n)
    xbar + c(-2, 2) * se
  } else if (type == "two.means") {
    if(!is.numeric(x) || !is.numeric(y)) {
      stop("Both x and y must be numeric values!")
    }
    xbar <- mean(x, ...)
    ybar <- mean(y, ...)
  
    estimate <- xbar - ybar
  
    n_x <- sum(!is.na(x))
    n_y <- sum(!is.na(y))
    s_x <- sd(x, ...)
    s_y <- sd(y, ...)
  
    se <- sqrt(s_x^2/n_x + s_y^2/n_y)
             
    estimate + c(-2, 2) * se
  } else {
    "Choose a type of either 'one.mean' or 'two.means'"
  }
}


```

(2) Write a function called `prop_ci()` for computing either a 95% confidence interval for a single proportion or a 95% confidence interval for the difference in proportions.

```{r}
prop_ci <- function(x, y=NULL, type="one.prop", ...) {
  if(type == "one.prop") {
    if(!is.numeric(x)) {
      stop("Please enter a numeric value for x!")
    } 
    for (num in 1:length(x)) {
     if(all(!is.na(x[num]), x[num] != 1, x[num] !=0)) {
        stop("Please only enter a 0/1 vector for x!")
         }
      }
    n.x <- length(x[!is.na(x)])
    p.x <- sum(x,...) / n.x
    estimate <- p.x
    se <- sqrt(p.x *(1 - p.x)/ n.x)
    
    estimate + c(-2,2) * se
  } else if (type == "two.props") {
    if(!is.numeric(x) || !is.numeric(y)) {
      stop("Both x and y must be numeric values!")
    }
    for (num in 1:length(x)) {
     if(all(!is.na(x[num]), x[num] != 1, x[num] !=0)) {
        stop("Please only enter a 0/1 vector for x!")
         }
      }
    for (num in 1:length(y)) {
     if(all(!is.na(y[num]), y[num] != 1, y[num] !=0)) {
        stop("Please only enter a 0/1 vector for y!")
         }
    }
    
    n.x <- length(x[!is.na(x)])
    p.x <- sum(x,...)/n.x

    n.y <- length(y[!is.na(y)])
    p.y <- sum(y,...)/n.y
    estimate <- p.x - p.y
    se <- sqrt(p.x*(1-p.x)/n.x + p.y*(1-p.y)/n.y)

    estimate + c(-2, 2) * se
  } else {
    "Choose a type of either 'one.prop' or 'two.props'"
  }
}



```


